On-the-fly Algorithms for GKAT Equivalences \thanks{Identify applicable funding agency here. If none, delete this.}

### 2025-01-16 main.tex
epsilon collapsing optimization

Another way to construct a coalgebra from a GKAT expression is via Thompson's construction, we lift the original construction to the symbolic setting.
A useful operation for Thompson's construction is the following guard operation, denoted by \(⟨B|\), where \(B\) is a set of boolean expressions: for a transition function \(δ̂\), a accepting function \(ϵ̂\), and a state \(s\),
\begin{align*}
    ⟨B|ϵ̂(s) & ≜ \{b ∧ c ∣ b ∈ B, c ∈ ϵ(s)\}; \\
    ⟨B|δ̂(s) & ≜ \{(b ∧ c, s', p) ∣ b ∈ B, (c, s', p) ∈ δ(s)\}.
\end{align*}
Besides guarding transition and acceptance with different conditions in if statements and while loops, the guard operator can also be used to simulate uniform continuation. Specifically, we can use \(⟨ϵ̂(s)|δ(s')\) to connecting all the accepting transition of \(s\) to the dynamic \(δ(s')\).

@@

>>>>

Another way to construct a coalgebra from a GKAT expression is via Thompson's construction, we lift the original construction to the symbolic setting.
A useful operation for Thompson's construction is the following guard operation, denoted by \(⟨B|\), where \(B\) is a set of boolean expressions: for a transition function \(δ̂\), a accepting function \(ϵ̂\), and a state \(s\),
\begin{align*}
    ⟨B|ϵ̂(s) & ≜ \{b ∧ c ∣ b ∈ B, c ∈ \hat{\epsilon}(s)\}; \\
    ⟨B|δ̂(s) & ≜ \{(b ∧ c, s', p) ∣ b ∈ B, (c, s', p) ∈ \hat{\delta}(s)\}.
\end{align*}
Besides guarding transition and acceptance with different conditions in if statements and while loops, the guard operator can also be used to simulate uniform continuation. Specifically, we can use \(⟨ϵ̂(s)|δ(s')\) to connecting all the accepting transition of \(s\) to the dynamic \(δ(s')\).

@@

The final optimization that we perform is collapsing the sets of boolean expressions of \(\hat{\epsilon}(s)\) into a single boolean expression. Basically, each \(\hat{\epsilon}(s)\) is represented by the boolean disjunction of its elements. Consider the guard operation defined in \Cref{sec:symb-gkat-construction}.
\begin{align*}
    ⟨\hat{\epsilon_1}(s_1)|\hat{\epsilon_2}(s_2) & ≜ \{b ∧ c ∣ b ∈ \hat{\epsilon_1}(s_1), c ∈ \hat{\epsilon_2}(s_2)\}
\end{align*}
If \(\hat{\epsilon_1}(s_1)\) and \(\hat{\epsilon_2}(s_2)\) are sets of boolean expressions, then the guard operation would produce a set whose size is proportional to \(|\hat{\epsilon_1}(s_1)| \times |\hat{\epsilon_2}(s_2)|\). However, if the collapsed representation is used, \(\hat{\epsilon_1}(s_1)\) and \(\hat{\epsilon_2}(s_2)\) become just boolean expressions instead of sets of expressions. The guard operation now produces a single boolean expression. The distributivity of boolean conjunction and disjunction ensures the correctness of this optimization.
